---
title: "Examining Affairs: an Exploratory Data Analysis Project"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, message = FALSE, warning = FALSE)
```

This is an examination of Infidelity data, known as Fair's Affairs: a cross-section dataset from a survey conducted by Psychology Today in 1969.


```{r, echo=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(AER)
library(wesanderson)
library(sjPlot)
library(knitr)

data("Affairs")

Affairs <- as_tibble(Affairs)

```


An initial look: how are affairs distributed? Does gender play a role?

```{r}
Affairs %>% 
  ggplot(aes(affairs))+
  geom_bar(aes(fill = gender))+
  scale_fill_manual(values=wes_palette(n=2, name = "GrandBudapest2"))
```

Let's introduce a new variable, faithful, for people who have never had an affair.
A binary variation of the variable, faithful_num, will be useful for visualising proportions.

```{r}
Affairs <- Affairs %>% 
  mutate(
    faithful = factor(ifelse(Affairs$affairs == 0, "Yes", "No"), levels = c("No", "Yes")),
    faithful_num = ifelse(Affairs$affairs == 0, 1, 0)
  ) 
```

## Comparing the proportion of faithful indviduals by gender:

```{r}
Affairs%>%
  group_by(gender) %>% 
  summarise(faithful_prop = mean(faithful_num)) %>% 
  ggplot(aes(gender, faithful_prop, fill = gender))+
  geom_col()+
  scale_fill_manual(values=wes_palette(n=2, name="GrandBudapest2"))+
  ylab("Proportion of faithful individuals")
```

Women appear to have a slightly higher proportion of faithfulness than men.

```{r}
tab_model(lm(faithful_num ~ gender, data = Affairs), p.style = "stars")
```

However, if we regress faithfulness against only gender, the significance appears very low.
Also, the R-squared and adjusted R- squared are both very small.

## What is the overall count of faithful vs non-faithful individuals?

```{r}
table(Affairs$faithful)
```

What is that proportionatelly?

```{r}
faithful_perc <- paste(round(100*mean(Affairs$faithful_num),2), "%")
(faithful_perc)
```

Okay, so we still have some faith in humanity left.

## Does having children impact faithfulness?

Let's peek at the overall distribution.

```{r}
Affairs %>% 
  ggplot(aes(children, affairs))+
  geom_count(aes(color = children))+
  scale_color_manual(values = wes_palette(n=2, name = "GrandBudapest2"))
```

A bar chart:

```{r}
Affairs%>% 
  ggplot(aes(children, fill = faithful))+
  geom_bar()+
  scale_fill_manual(values=wes_palette(n=2, name="GrandBudapest2"))
```

How does faithfulness look proportionatelly?

```{r}
Affairs %>% 
  group_by(children) %>% 
  summarise(faith_prop = mean(faithful_num))
```

A graph with proportions:

```{r}
Affairs %>% 
  group_by(children) %>% 
  summarise(faith_prop = mean(faithful_num)) %>% 
  ggplot(aes(children, faith_prop, fill = children))+
  geom_col()+
  scale_fill_manual(values = wes_palette(n=2, name = "GrandBudapest2"))+
  ylab("Proportion of faithful individuals")
```

So, people without children appear to have less affairs.

## Are unhappy couples more likely to be unfaithful?

Let's make life easier for ourselves by updating "rating" and "religiousness" to factors.

```{r}
Affairs <- Affairs %>% 
  mutate(
    rating =  factor(rating, levels = c("1", "2", "3", "4", "5")),
    religiousness =  factor(religiousness, levels = c("1", "2", "3", "4", "5"))
  )
```

Now a simple look at the overall distribution:

```{r}
Affairs %>% 
  ggplot(aes(rating, affairs))+
  geom_col(aes(fill = rating))+
  scale_fill_manual(values = wes_palette(n = 5, name = "Moonrise3"))+
  xlab("Happiness rating")
```

Weird, could unhappy couples really be less prone to affairs than very happy ones?
Let's look at faithfulness proportions instead.

```{r}
Affairs %>% 
  group_by(rating) %>% 
  summarise(faith_prop = mean(faithful_num)) %>% 
  ggplot(aes(rating, faith_prop))+
  geom_col(aes(fill = rating))+
  scale_fill_manual(values = wes_palette(n = 5, name = "GrandBudapest2", type = "continuous"))+
  xlab("Happiness rating")+
   ylab("Proportion of faithful individuals")
```

Okay, so when we look at faithfulness proportionatelly, it matches the common sense assumption that happier couples will be less prone to cheating.

## What about religiousness, does it have an impact?

An initial look:

```{r}
Affairs %>% 
  ggplot(aes(religiousness, affairs))+
  geom_col(aes(fill = religiousness))+
  scale_fill_manual(values = wes_palette(n = 5, name = "Rushmore1", type = "continuous"))
```

Faithfulness proportionatelly:

```{r}
Affairs %>% 
  group_by(religiousness) %>%
  summarise(faith_prop = mean(faithful_num)) %>% 
  ggplot()+
  geom_col(aes(religiousness,faith_prop, fill = religiousness))+
  scale_fill_manual(values = wes_palette(n = 5, name = "Rushmore1", type = "continuous"))+
  ylab("Proportion of faithful individuals")
```

Anti-religous people seem to have a lower proportion of faithfulness than others.

## Is there a link between religiousness and having children?

Let's make a binary variable for having children.

```{r}
Affairs %>% 
  mutate(
    children_bin = ifelse(children == "yes", 1, 0)
  ) %>% 
  group_by(religiousness) %>% 
  summarise(child_proportion = mean(children_bin))
```

A graph:

```{r}
Affairs %>% 
  mutate(
    children_bin = ifelse(children == "yes", 1, 0)
  ) %>% 
  group_by(religiousness) %>% 
  summarise(child_proportion = mean(children_bin)) %>% 
  ggplot(aes(religiousness, child_proportion))+
  geom_col(aes(fill = religiousness))+
  scale_fill_manual(values = wes_palette(n = 5, name = "Rushmore1", type = "continuous"))+
  ylab("Proportion of individuals with children")
```

There appears to be a link.

## What about education and faithfulness?

Let's create a new facor with names and concatenate "some college" with "high school graduate" and "some graduate work" with "college graduate", as these two levels seem arbitrary.

```{r}
Affairs <- Affairs %>% 
  mutate(
    education_ft = factor(education, levels = c("9", "12", "14", "16", "17", "18", "20")),
    education_ft = fct_recode(education_ft,
                              "grade_school" = "9",
                              "high_school" = "12",
                              "high_school" = "14",
                              "college" = "16",
                              "college" = "17",
                              "masters" = "18",
                              "phd_or_similar" = "20"
    )
  )
```

Looking into faithfulness proportions again.

```{r}
Affairs %>% 
  group_by(education_ft) %>% 
  summarise(faithful_prop = mean(faithful_num)) %>% 
  ggplot(aes(education_ft, faithful_prop))+
  geom_col(aes(fill = education_ft))+
  scale_fill_manual(values = wes_palette(n = 5, name = "GrandBudapest1", type = "continuous"))+
  ylab("Proportion of faithful individuals")
```

There doesn't appear to be a clear connection.

## What about occupation?

```{r}
Affairs %>% 
  mutate(occupation = factor(occupation, levels = 1:7)) %>% 
  group_by(occupation) %>% 
  summarise(faithful_prop = mean(faithful_num))%>% 
  ggplot(aes(occupation, faithful_prop))+
  geom_col(aes(fill = occupation))+
  scale_fill_manual(values = wes_palette(n = 7, name = "Royal2", type = "continuous"))+
  ylab("Proportion of faithful individuals")
```

Again, no clear connection.

## Linear regressions 

```{r}
lm1 <- lm(affairs ~ gender + children + religiousness + rating + yearsmarried + age , data = Affairs)
tab_model(lm1, p.style = "stars")
```

Does including education change anything?

```{r}
lm2 <- lm(affairs ~ gender + children + religiousness + rating + yearsmarried + age + education, data = Affairs)
tab_model(lm2, p.style = "stars")
```

Adjusted R-squared is now lower, which indicates a lower accuracy than before.

## What about including occupation?

```{r}
lm3 <- lm(affairs ~ gender + children + religiousness + rating + yearsmarried + age + occupation, data = Affairs)
tab_model(lm3, p.style = "stars")
```

Adjusted R-squared has increased, but occupation is not marked as significant.

## How about including both?

```{r}
lm4 <- lm(affairs ~ gender + children + religiousness + rating + yearsmarried + age + occupation + education, data = Affairs)
tab_model(lm4, p.style = "stars")
```

The adjusted R-squared is, again, lower and the two variables do not appear to be that significant.


## The linear model from the documentation:

```{r}
fm_ols <- lm(affairs ~ age + yearsmarried + religiousness + occupation + rating,
             data = Affairs)

tab_model(fm_ols, p.style = "stars")
```

It seems like removing the "children" variable brings more accuracy to the model.

